
```{r}
library(tidyverse)
library(gridExtra) # for arranging multiple plots in 1 image
```


### [1] DATA WRANGLING

Combine data from 2 observers and sample info from master sheet
```{r}
raw_pigmentation1 <- read.csv("./data/Caitlynn_PigmentationData.csv")
raw_pigmentation2 <- read.csv("./data/Andy_PigmentationData.csv")
pigmentation_master <- read.csv("./data/Samples_Master.csv")

# Combine data and recompute sums
raw_pigmentation1 <- raw_pigmentation1 %>% rename(Rand_ID = Tube.)
pigmentation_wTPs28 <- rbind(raw_pigmentation1[1:5], raw_pigmentation2[1:5])
pigmentation_wTPs28 <- pigmentation_wTPs28 %>% mutate("Sum" = T5 + T6 + T7)

# Merge using Rand_ID
pigmentation_wTPs28 <- merge(pigmentation_wTPs28, pigmentation_master)
write.csv(pigmentation_wTPs28, "pigmentation_wTPs28.csv")

# Notes:
# - raw_pigmentation1 row 1557 malformed
# - no data for Rand_ID 13
```

Remove time points 2 and 8 from data
```{r}
pigmentation <- pigmentation_wTPs28[!(pigmentation_wTPs28$TP==2 |pigmentation_wTPs28$TP==8),]
write.csv(pigmentation, "pigmentation.csv")
```

Create data frame with 1 row per sample
```{r}
# pigmentation <- read.csv("pigmentation.csv")

pigmentation_nadrop <- drop_na(pigmentation) # TEMP CODE, Removed NAs to visualize
pigmentation_by_TP_cage <- pigmentation_nadrop %>% # TEMP CODE, Removed NAs to visualize
# pigmentation_by_TP_cage <- pigmentation %>% 
  group_by(TP, Treatment, Cage) %>% 
  summarise(
    T5 = mean(T5),
    T6 = mean(T6),
    T7 = mean(T7),
    Sum = mean(Sum)
  ) %>% 
  mutate(
    Cage = paste(Treatment, Cage, sep="-"),
    TP = as.factor(TP),
    Treatment = case_when(
      Treatment == "E" ~ "Control",
      Treatment == "P" ~ "Experimental"
    )
  ) 

write.csv(pigmentation_by_TP_cage, "pigmentation_by_TP_cage_nadrop.csv") # TEMP CODE, Removed NAs to visualize
# write.csv(pigmentation_by_TP_cage, "pigmentation_by_TP_cage.csv")

summary(pigmentation_by_TP_cage[4:7])
```

Create file with 1 row per TP / Treatment
```{r}
# pigmentation <- read.csv("pigmentation.csv")

pigmentation_nadrop <- drop_na(pigmentation) # TEMP CODE, Removed NAs to visualize
pigmentation_by_TP_treatment <- pigmentation_nadrop %>% # TEMP CODE, Removed NAs to visualize
# pigmentation_by_TP_treatment <- pigmentation %>% 
  group_by(TP, Treatment) %>%
  summarise(
    se5 = sd(T5)/sqrt(10),
    se6 = sd(T6)/sqrt(10),
    se7 = sd(T7)/sqrt(10),
    seSum = sd(Sum)/sqrt(10),
    T5 = mean(T5),
    T6 = mean(T6),
    T7 = mean(T7),
    Sum = mean(Sum),
    ) %>% 
    mutate(
      TP = as.factor(TP),
      Treatment = case_when(
        Treatment == "E" ~ "Control",
        Treatment == "P" ~ "Experimental"
      )
    )

write.csv(pigmentation_by_TP_treatment, "pigmentation_by_TP_treatment_nadrop.csv") # TEMP CODE
# write.csv(pigmentation_by_TP_Treatment, "pigmentation_by_TP_Treatment.csv")

summary(pigmentation_by_TP_cage[4:7])
```


### [2] VISUALIZATION ### 

Read in data (if did not run code above)
```{r}
# Use these .csv files if data frames have not been fixed
# pigmentation_by_TP_treatment <- read.csv("pigmentation_by_TP_cage_nadrop.csv")
# pigmentation_by_TP_treatment <- read.csv("pigmentation_by_TP_treatment_nadrop.csv")

# pigmentation_by_TP_treatment <- read.csv("pigmentation_by_TP_cage.csv")
# pigmentation_by_TP_treatment <- read.csv("pigmentation_by_TP_treatment.csv")
```

Pigmentation visualizations
```{r}
p1 <- ggplot(
  data=pigmentation_by_TP_cage,
  aes(x=TP, y=T5, group=Treatment, color=Treatment)
) + 
geom_line(aes(group=Cage), alpha=0.2) + # Cage pigmentation
geom_point(aes(color=Treatment), size= 0.2, alpha = 0.2) +  # Cage pigmentation
geom_line(data=pigmentation_by_TP_treatment, aes(x=TP, y=T5), size=1) +  # Treatment pigmentation
geom_errorbar(data=pigmentation_by_TP_treatment, aes(y=T5, ymin=T5-se5, ymax=T5+se5), width=.2,
               position=position_dodge(0.05)) + # Treatment pigmentation
scale_color_manual(values=c("black","orange")) +
scale_x_discrete(
  labels = c("Jul 26",  "Aug 25", "Sep 7", "Dec 22"),
) +
ylim(0, 10) +
labs( y = "T5 Pigmentation Score",  x="") +
theme_classic() # +
# theme(
#   axis.title.x = element_text(size = 0),
#   axis.title.y = element_text(size = 30),
#   axis.text.x = element_text(size = 30),
#   axis.text.y = element_text(size = 30),
#   )  

p2 <- ggplot(
  data=pigmentation_by_TP_cage,
  aes(x=TP, y=T6, group=Treatment, color=Treatment)
) + 
geom_line(aes(group=Cage), alpha=0.2) + # Cage pigmentation
geom_point(aes(color=Treatment), size= 0.2, alpha = 0.2) + # Cage pigmentation
geom_line(data=pigmentation_by_TP_treatment, aes(x=TP, y=T6), size=1) + # Treatment pigmentation
geom_errorbar(data=pigmentation_by_TP_treatment, aes(y=T6, ymin=T6-se6, ymax=T6+se6), width=.2,
               position=position_dodge(0.05)) + # Treatment pigmentation
scale_color_manual(values=c("black","orange")) +
scale_x_discrete(
  labels = c("Jul 26",  "Aug 25", "Sep 7", "Dec 22"),
) +
ylim(0, 10) +
labs( y = "T6 Pigmentation Score",  x="") +
theme_classic() # +
# theme(
#   axis.title.x = element_text(size = 0),
#   axis.title.y = element_text(size = 30),
#   axis.text.x = element_text(size = 30),
#   axis.text.y = element_text(size = 30),
#   )  

p3 <- ggplot(
  data=pigmentation_by_TP_cage,
  aes(x=TP, y=T7, group=Treatment, color=Treatment)
) + 
geom_line(aes(group=Cage), alpha=0.2) + # Cage pigmentation
geom_point(aes(color=Treatment), size= 0.2, alpha = 0.2) + # Cage pigmentation
geom_line(data=pigmentation_by_TP_treatment, aes(x=TP, y=T7), size=1) + # Treatment pigmentation
geom_errorbar(data=pigmentation_by_TP_treatment, aes(y=T7, ymin=T7-se7, ymax=T7+se7), width=.2,
               position=position_dodge(0.05)) + # Treatment pigmentation
scale_color_manual(values=c("black","orange")) +
scale_x_discrete(
  labels = c("Jul 26",  "Aug 25", "Sep 7", "Dec 22"),
) +
ylim(0, 10) +
labs( y = "T7 Pigmentation Score",  x="") +
theme_classic() # +
# theme(
#   axis.title.x = element_text(size = 0),
#   axis.title.y = element_text(size = 30),
#   axis.text.x = element_text(size = 30),
#   axis.text.y = element_text(size = 30),
#   )  

p4 <- ggplot(
  data=pigmentation_by_TP_cage,
  aes(x=TP, y=Sum, group=Treatment, color=Treatment)
) + 
geom_line(aes(group=Cage), alpha=0.2) + # Cage pigmentation
geom_point(aes(color=Treatment), size= 0.2, alpha = 0.2) + # Cage pigmentation
geom_line(data=pigmentation_by_TP_treatment, aes(x=TP, y=Sum), size=1) +
geom_point(data=pigmentation_by_TP_treatment, aes(x=TP, y=Sum)) + # Treatment pigmentation
geom_errorbar(data=pigmentation_by_TP_treatment, aes(y=Sum, ymin=Sum-seSum, ymax=Sum+seSum), width=.2,     
              position=position_dodge(0.05)) + # Treatment pigmentation
scale_color_manual(values=c("black","orange")) +
scale_x_discrete(
  labels = c("Jul 26",  "Aug 25", "Sep 7", "Dec 22"),
) +
labs(y = "Sum Pigmentation Score",  x="") +
theme_classic() # +
# theme(
#   axis.title.x = element_text(size = 0),
#   axis.title.y = element_text(size = 28),
#   axis.text.x = element_text(size = 30),
#   axis.text.y = element_text(size = 30),
#   legend.title = element_text(size = 30),
#   legend.text = element_text(size = 25)
#   )  

grid.arrange(p1, p2, p3, p4)
ggsave("pigmentation_plots.png") # Only saves sum plot
```


### [3] STATISTICS ###

Read in data (if did not run code above)
```{r}
# pigmentation <- read.csv("pigmentation.csv")
# pigmentation_nadrop <- drop_na(pigmentation)
```

Normality test
```{r}
shapiro.test(pigmentation_nadrop$Sum)

boxplot(pigmentation_nadrop$Sum)
boxplot(pigmentation_nadrop$Sum)
```

Create new data frame from pigmentation_nadrop with log(Sum) because data is NOT normal
```{r}
pigmentation_nadrop_log <- pigmentation_nadrop %>% mutate(log.Sum = log(Sum))
```

Analyzing time using linear mixed effect model (for each time frame)
```{r}
stats_time_pigmentation <- data.frame(treatment=character(), time_frame=integer(), p=double())

for (group in unique(pigmentation_nadrop_log$Treatment)) {
  time_gap = 1
  for (dates in list(c(3, 7), c(7, 9), c(9, 13))) {
    data_sub <- pigmentation_nadrop_log %>% filter(TP %in% dates & Treatment == group)
    
    full.model = lmer(log.Sum ~ TP + (1|Experimenter) + (1|Sample), data=data_sub, REML=FALSE)
    noTP.model = lmer(log.Sum ~ (1|Experimenter) + (1|Sample), data=data_sub, REML=FALSE)
    p = anova(full.model, noTP.model)[2,8]
    
    stats_time_pigmentation  <- rbind(stats_time_pigmentation, data.frame(treatment=c(group), time_frame=c(time_gap), p=c(p)))
    time_gap = time_gap + 1
  }
}

# Make easier to interpret
stats_time_pigmentation_simple <- stats_time_pigmentation  %>%
  mutate(
    time_frame = case_when(
      # time_frame == 1 ~ "Jul 26 - Aug 25",
      # time_frame == 2 ~ "Aug 25 - Sep 7",
      # time_frame == 3 ~ "Sep 7 - Dec 22"
      time_frame == 1 ~ "07-26 to 08-25",
      time_frame == 2 ~ "08-25 to 09-07",
      time_frame == 3 ~ "09-07 to 12-22"
    ),
    p = case_when(
      p < 0.001 ~ "<0.001",
      p >= 0.001 ~ as.character(round(p, digits=3))
    ),
    sig = case_when(
      p < 0.05 ~ "*",
      p >= 0.001 ~ ""
    )
  )

write.csv(stats_time_pigmentation_simple, "stats_time_pigmentation.csv")
stats_time_pigmentation_simple
```

Analyzing treatment using linear mixed effect model (at each date)
```{r}
stats_treatment_pigmentation <- data.frame(TP=integer(), p=double())

for (tp in c(3, 7, 9, 13)) {
  data_sub <- pigmentation_nadrop_log %>% filter(TP == tp)
  full.model = lmer(log.Sum ~ Treatment + (1|Experimenter) + (1|Sample), data=data_sub, REML=FALSE)
  noTreatment.model = lmer(log.Sum ~ (1|Experimenter) + (1|Sample), data=data_sub, REML=FALSE)
  p = anova(full.model, noTreatment.model)[2,8]
  stats_treatment_pigmentation <- rbind(stats_treatment_pigmentation, data.frame(TP=c(tp), p=c(p)))
}

# Make easier to interpret
stats_treatment_pigmentation_simple <- stats_treatment_pigmentation %>%
  mutate(
    date = case_when(
      TP == 3 ~ "07-26",
      TP == 7 ~ "08-25",
      TP == 9 ~ "09-07",
      TP == 13 ~ "12-22"
      # TP == 3 ~ "Jul 26",
      # TP == 7 ~ "Aug 25",
      # TP == 9 ~ "Sep 7",
      # TP == 13 ~ "Dec 22"
    ),
    p = case_when(
      p < 0.001 ~ "<0.001",
      p >= 0.001 ~ as.character(round(p, digits=3))
    ),
    sig = case_when(
      p < 0.05 ~ "*",
      p >= 0.001 ~ ""
    )
  ) %>% 
  select(date, p, sig)

write.csv(stats_treatment_pigmentation_simple, "stats_treatment_pigmentation.csv")
stats_treatment_pigmentation_simple 
```
