
```{r}
library(tidyverse)
library(lme4) # for modeling
```


### [1] DATA WRANGLING ###

Combine all .csv files for each into 1 data frame 
```{r}
# INPUT DATA: 7 .csv file, each for an image folder (columns: image, count)
data_dir = "./data/" # Combines all .csv files in specified folder

read_plus <- function(flnm) {
    read_csv(flnm) %>% 
        mutate(filename = flnm)
}

raw_data <-
    list.files(path = data_dir,
               pattern = "*.csv", 
               full.names = T) %>% 
    map_df(~read_plus(.))
```

Reorganize data frame for relevant info 
OUTPUT DATA: (columns: date, treatment, cage, replicate, id, count)
```{r}
# INPUT DATA: (columns: image [includes cage], count, filename [includes date])
census <- raw_data %>% 
  mutate(
    date = substr(filename, 15, 19),
    treatment = substr(image, 1, 1),
    cage = as.integer(case_when( 
      nchar(image) == 4 ~ substr(image, 2, 2), # EX: "E1-A" -> "E1"
      nchar(image) == 5 ~ substr(image, 2, 3)  # EX: "E10-A" -> "E10"
    )),
    replicate  = case_when(
      nchar(image) == 4 ~ substr(image, 4, 4), # EX: "E1-A" -> "A"
      nchar(image) == 5 ~ substr(image, 5, 5)  # EX: "E10-A" -> "A"
    ),
  ) %>% 
  select(date, treatment, cage, replicate, count) %>% 
  arrange(date, treatment, cage, replicate) %>% 
  mutate(
    count = count * 40 # Because region of interest is 2.5% of the cage
  )

write.csv(census, "census.csv")
```

Mean replicate counts for cage census
OUTPUT DATA: (columns: date, treatment, cage, id, census)
```{r}
census_by_cage <- census %>% 
  group_by(date, treatment, cage) %>% 
  summarise(census = mean(count)) %>% # Mean 4 cage replicates
  mutate(
    id = paste(treatment, cage, sep="-"), # id: Unique treatment/cage identifier
    treatment = case_when( # Rename treatments
      treatment == "E" ~ "Control",
      treatment == "P" ~ "Experimental"
    )
  )

write.csv(census_by_cage, "census_by_cage.csv")
```

Mean replicate counts for treatment census
OUTPUT DATA: (columns: date, treatment, cage, id, census)
```{r}
census_by_treatment <- census %>% 
  group_by(date, treatment) %>% 
  summarise(
    census = mean(count), # Mean all treatment counts
    se_census = sd(count) / sqrt(4)
  ) %>% 
  mutate(
    treatment = case_when( # Rename treatments
      treatment == "E" ~ "Control",
      treatment == "P" ~ "Experimental"
    )
  )

write.csv(census_by_treatment, "census_by_treatment.csv")
```


### [2] VISUALIZATION ###

Plot census over time for each cage and for each treatment
```{r}
ggplot(
  data=census_by_cage,
  aes(x=date, y=census, group=treatment, color=treatment)
) + 
geom_line(aes(group=id), alpha=0.2) + # Cage census
geom_point(aes(color=treatment), size= 0.2, alpha = 0.2) + # Cage census
geom_line(data=census_by_treatment, aes(x=date, y=census), size=1) + # Mean census
geom_point(data=census_by_treatment, aes(x=date, y=census)) + # Mean census
geom_errorbar(data=census_by_treatment, # Mean census
              aes(x=date, y=census, ymin=census-se_census, ymax=census+se_census),
              width=.2, position=position_dodge(0.05)) +
scale_color_manual(values=c("black","orange")) +
scale_x_discrete(
  labels = c("Jul 19",  "Aug 5", "Aug 20", "Sep 17"),
) +
labs(y = "Adult Census Estimate",  x="", colour="Treatment") +
theme_classic() # +
# theme(
#   axis.title.x = element_text(size = 0),
#   axis.title.y = element_text(size = 30),
#   axis.text.x = element_text(size = 30),
#   axis.text.y = element_text(size = 30),
#   legend.title = element_text(size = 30),
#   legend.text = element_text(size = 25)
#   )

ggsave("census_plot.png")
# ggsave("census_plot.png", width = 20, height = 20, units = "cm")
```


### [3] STATISTICS ###

Normality test
```{r}
shapiro.test(census_by_cage$census)

boxplot(census$count)
boxplot(census_by_cage$census)
```

Create new data frame from census with log(census) because data is NOT normal
```{r}
census_log <- census %>%
  mutate(
    log.census = case_when(
      count == 0 ~ 0,
      TRUE ~ log(count * 40),
    ),
    id = paste(treatment, cage, sep="-")
  )
```

Analyzing time using linear mixed effect model (for each time frame)
```{r}
stats_time_census <- data.frame(treatment=character(), time_frame=integer(), p=double())

for (group in unique(census_log$treatment)) {
  for (dates in list(c("07-19", "08-05"), c("08-05", "08-20"), c("08-20", "09-17"))) {
    data_sub <- census_log %>% filter(date %in% dates, treatment == group)
    
    full.model = lmer(log.census ~ date + (1|id), data=data_sub, REML=FALSE)
    noTP.model = lmer(log.census ~ (1|id), data=data_sub, REML=FALSE)
    p = anova(full.model, noTP.model)[2,8]
    
    time_gap = paste(dates[1], dates[2], sep=" to ")
    stats_time_census <- rbind(stats_time_census, data.frame(treatment=c(group), time_frame=c(time_gap), p=c(p)))
  }
}

# Make easier to interpret
stats_time_census_simple <- stats_time_census %>%
  mutate(
    p = case_when(
      p < 0.001 ~ "<0.001",
      p >= 0.001 ~ as.character(round(p, digits=3))
    ),
    sig = case_when(
      p < 0.05 ~ "*",
      TRUE ~ ""
    )
  )

stats_time_census_simple
```

Analyzing treatment using linear mixed effect model (at each date)
```{r}
stats_treatment_census <- data.frame(date=character(), p=double())

for (day in c("07-19", "08-05", "08-20", "09-17")) {
  data_sub <- census_log %>% filter(date == day)
  
  full.model = lmer(log.census ~ treatment + (1|id), data=data_sub, REML=FALSE)
  noTreatment.model = lmer(log.census ~ (1|id), data=data_sub, REML=FALSE)
  p = anova(full.model, noTreatment.model)[2,8]
  
  stats_treatment_census <- rbind(stats_treatment_census, data.frame(date=c(day), p=c(p)))
}

# Make easier to interpret
stats_treatment_census_simple <- stats_treatment_census %>%
  mutate(
    p = case_when(
      p < 0.001 ~ "<0.001",
      p >= 0.001 ~ as.character(round(p, digits=3))
    ),
    sig = case_when(
      p < 0.05 ~ "*",
      TRUE ~ ""
    )
  )

stats_treatment_census_simple
```
