
```{r}
library(tidyverse)
```


### [1] DATA WRANGLING ###

Snippet reads and combines all .csv files in ./data/ subdirectory
```{r}
read_plus <- function(flnm) {
    read_csv(flnm) %>% 
        mutate(filename = flnm)
}

raw_data <-
    list.files(path = "./data/",
               pattern = "*.csv", 
               full.names = T) %>% 
    map_df(~read_plus(.))
```

Clean into data frame with columns: date, treatment, cage, replicate, count
```{r}
# data currently has columns: image (includes cage), count, ..., filename (includes date)
census <- raw_data %>% 
  mutate(
    date = substr(filename, 15, 19),
    treatment = substr(image, 1, 1),
    cage = as.integer(case_when(
      nchar(image) == 4 ~ substr(image, 2, 2),
      nchar(image) == 5 ~ substr(image, 2, 3)
    )),
    replicate  = case_when(
      nchar(image) == 4 ~ substr(image, 4, 4),
      nchar(image) == 5 ~ substr(image, 5, 5)
    ),
  ) %>% 
  select(date, treatment, cage, replicate, count) %>% 
  arrange(date, treatment, cage, replicate) %>% 
  mutate(
    id = paste(treatment, cage, sep="-"),
    count = count * 40
    )
```

Summarise into data frame with columns: date, treatment, cage, count (average replicates * 40)
```{r}
census_by_cage <- census %>% 
  group_by(date, treatment, cage) %>% 
  summarise(population = mean(count)) %>% 
  mutate(
    id = paste(treatment, cage, sep="-"),
    treatment = case_when(
      treatment == "E" ~ "Control",
      treatment == "P" ~ "Experimental"
    )
  )
```

Summarise into data frame with columns: date, treatment, count (average replicates)
```{r}
census_by_treatment <- census %>% 
  group_by(date, treatment) %>% 
  summarise(
    population = mean(count),
    se_population = sd(count) / sqrt(4)
  ) %>% 
  mutate(
    treatment = case_when(
      treatment == "E" ~ "Control",
      treatment == "P" ~ "Experimental"
    )
  )
```



### [2] DATA VISUALIZATION ###

```{r}
plotCensus <- ggplot(
  data=census_by_cage,
  aes(x=date, y=population, group=treatment, color=treatment)
) + 
geom_line(aes(group=id), alpha=0.2) +
geom_point(aes(color=treatment), size= 0.2, alpha = 0.2) + 
geom_line(data=census_by_treatment, aes(x=date, y=population), size=1) +
geom_point(data=census_by_treatment, aes(x=date, y=population)) + 
geom_errorbar(data=census_by_treatment,
              aes(x=date, y=population, ymin=population-se_population, ymax=population+se_population),
              width=.2, position=position_dodge(0.05))+
scale_color_manual(values=c("black","orange"))+
scale_x_discrete(
  labels = c("Jul 19",  "Aug 5", "Aug 20", "Sep 17"),
) +
#ylim(0, 10) +
labs(y = "Adult Census Estimate",  x="", colour="Treatment") +
theme_classic() +
theme(
  axis.title.x = element_text(size = 0),
  axis.title.y = element_text(size = 30),
  axis.text.x = element_text(size = 30),
  axis.text.y = element_text(size = 30),
  legend.title = element_text(size = 30),
  legend.text = element_text(size = 25)
  )  
```



### [3] STATISTICS ###

```{r}
library(lme4)
library(pixiedust)
library(kableExtra)
```

```{r}
shapiro.test(census_by_cage$population)

boxplot(census$count)
boxplot(census_by_cage$population)
```

```{r}
census_log <- census %>%
  mutate(
    log.population = case_when(
      count == 0 ~ 0,
      TRUE ~ log(count * 40)
    )
  )
```

```{r}
stats_time_census <- data.frame(treatment=character(), time_gap=integer(), p=double())

for (t in unique(census_log$treatment)) {
  for (time_frame in list(c("07-19", "08-05"), c("08-05", "08-20"), c("08-20", "09-17"))) {
    data_sub <- census_log %>% filter(date %in% time_frame, treatment == t)
    full.model = lmer(log.population ~ date + (1|id), data=data_sub, REML=FALSE)
    noTP.model = lmer(log.population ~ (1|id), data=data_sub, REML=FALSE)
    p = anova(full.model, noTP.model)[2,8]
    time_gap = paste(time_frame[1], time_frame[2], sep=" to ")
    stats_time_census <- rbind(stats_time_census, data.frame(treatment=c(t), time_gap=c(time_gap), p=c(p)))
  }
}

stats_time_census2 <- stats_time_census %>%
  mutate(
    p = case_when(
      p < 0.001 ~ "<0.001",
      p >= 0.001 ~ as.character(round(p, digits=3))
    ),
    sig = case_when(
      p < 0.05 ~ "*",
      TRUE ~ ""
    )
  )
```

```{r}
stats_treatment_census <- data.frame(date=character(), p=double())

for (d in c("07-19", "08-05", "08-20", "09-17")) {
  data_sub <- census_log %>% filter(date == d)
  full.model = lmer(log.population ~ treatment + (1|id), data=data_sub, REML=FALSE)
  noTreatment.model = lmer(log.population ~ (1|id), data=data_sub, REML=FALSE)
  p = anova(full.model, noTreatment.model)[2,8]
  stats_treatment_census <- rbind(stats_treatment_census, data.frame(TP=c(d), p=c(p)))
}

stats_treatment_census2 <- stats_treatment_census %>%
  mutate(
    p = case_when(
      p < 0.001 ~ "<0.001",
      p >= 0.001 ~ as.character(round(p, digits=3))
    ),
    sig = case_when(
      p < 0.05 ~ "*",
      TRUE ~ ""
    )
  )

```

















Analyzing time using linear mixed effect model
```{r}
census_by_cage_log <- census_by_cage %>%  mutate(log.population = log(population))

stats_time_census <- data.frame(treatment=character(), time_gap=integer(), p=double())

for (t in unique(census_by_cage_log$treatment)) {
  for (time_frame in list(c("07-19", "08-05"), c("08-05", "08-20"), c("08-20", "09-17"))) {
    data_sub <- census_by_cage_log %>% filter(date %in% time_frame, treatment == t)
    full.model = lmer(log.population ~ date + (1|id), data=data_sub, REML=FALSE)
    noTP.model = lmer(log.population ~ (1|id), data=data_sub, REML=FALSE)
    p = anova(full.model, noTP.model)[2,8]
    time_gap = paste(time_frame[1], time_frame[2], sep=" to ")
    stats_time_census <- rbind(stats_time_census, data.frame(treatment=c(t), time_gap=c(time_gap), p=c(p)))
  }
}

stats_time_census2 <- stats_time_census %>%
  mutate(
    p = case_when(
      p < 0.001 ~ "<0.001",
      p >= 0.001 ~ as.character(round(p, digits=3))
    ),
    sig = case_when(
      p < 0.05 ~ "*",
      TRUE ~ ""
    )
  )
```

```{r}
stats_treatment_census <- data.frame(date=character(), p=double())

for (d in c("07-19", "08-05", "08-20", "09-17")) {
  data_sub <- census_by_cage_log %>% filter(date == d)
  full.model = lmer(log.population ~ treatment + (1|id), data=data_sub, REML=FALSE)
  noTreatment.model = lmer(log.population ~ (1|id), data=data_sub, REML=FALSE)
  p = anova(full.model, noTreatment.model)[2,8]
  stats_treatment_census <- rbind(stats_treatment_census, data.frame(TP=c(tp), p=c(p)))
}

stats_treatment_census2 <- stats_treatment_census %>%
  mutate(
    p = case_when(
      p < 0.001 ~ "<0.001",
      p >= 0.001 ~ as.character(round(p, digits=3))
    ),
    sig = case_when(
      p < 0.05 ~ "*",
      TRUE ~ ""
    )
  )

```







